---
title: "rMATS Processing: Subsets rMATS output for rMATS-STAT"
output: html_document
date: "3/13/2019 edited 3/27/2019"
output:
  html_document:
    toc: true
    toc_depth: 3
---

This document creates in the input files for rMATS_STAT, the bashscripts joining multiple jobs for the HPC and the batchscripts needed to submit them to the HPC
This version of the script uses healthy bone marrow as the control group. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(data.table)
```


####Read in final mutations file
```{r}
#somatic mutations
mutations <- read.csv("../inputfiles/filtered_tidied/2019_03_01_combined_mutations_deletions_controls.txt", header = TRUE, sep="\t", stringsAsFactors = FALSE)
binary <- read.csv("../inputfiles/filtered_tidied/2019_03_01_combined_mutations_deletions_controls_binary.txt", header=TRUE, sep="\t", stringsAsFactors = FALSE)
ids <- read.csv("../inputfiles/filtered_tidied/2019_03_01_filtered_inclengths.txt", header = TRUE, sep="\t", stringsAsFactors = FALSE)
length(unique(mutations$array_id))
```

```{r}

binary_subset <- binary[,grep("PRPF8_D1598|DDX41_R525H",colnames(binary))]

#Add Healthy BM column
binary_subset$healthy_bm <- ifelse(rownames(binary_subset) %in% mutations[mutations$mkh=="healthy_bm","array_id"],"1","0")

k <- which(binary_subset>=1, arr.ind=TRUE)
binary_subset[k] <- rownames(k)

binary_subset[binary_subset == 0] <- NA
```

###Read in rMATS output
```{r}
filtered_IJC <- fread("../inputfiles/filtered_tidied/2019_03_01_filtered_includedjunctioncounts.txt", sep="\t", stringsAsFactors = FALSE, data.table=FALSE)
rownames(filtered_IJC) <- filtered_IJC$V1
filtered_IJC$V1 <- NULL

filtered_SJC <- fread("../inputfiles/filtered_tidied/2019_03_01_filtered_skippedjunctioncounts.txt",  sep="\t", stringsAsFactors = FALSE, data.table=FALSE)
rownames(filtered_SJC) <- filtered_SJC$V1
filtered_SJC$V1 <- NULL

inclengths <- fread("../inputfiles/filtered_tidied/2019_03_01_filtered_inclengths.txt", sep="\t", stringsAsFactors = FALSE, data.table=FALSE)
rownames(inclengths) <- inclengths$V1
inclengths$V1 <- NULL
```
###Remove MXE
```{r}
filtered_IJC <- filtered_IJC[which(inclengths$type!="MXE"),]
filtered_SJC <- filtered_SJC[which(inclengths$type!="MXE"),]
inclengths <- inclengths[which(inclengths$type!="MXE"),]
```

##Creating RMATS_STAT input divided by disease type and splicing factor mutation
These SF mutations have enough samples to be run for each disease
```{r}
make_rmats_stat_files <- function(column){

  #disease <- "MDS"
  test <- column
  
  condition <- binary_subset[,column]
  condition <- condition[!is.na(condition)]
  print(condition)
  
  control <- binary_subset$healthy_bm[!is.na(binary_subset$healthy_bm)]
  
  if(length(condition) >= 2 & length(control>= 2)){
    
    comparison <- paste(test,"_versus_","healthy_bm", sep="")   
    
    ##Subset the dataframe for samples of interest
        
    filtered_IJC_condition <- filtered_IJC[,which(colnames(filtered_IJC) %in% condition)]
    filtered_SJC_condition <- filtered_SJC[,which(colnames(filtered_SJC) %in% condition)]
        
    filtered_IJC_control <- filtered_IJC[,which(colnames(filtered_IJC) %in% control)]
    filtered_SJC_control <- filtered_SJC[,which(colnames(filtered_SJC) %in% control)]
        
    ###Create final dataframe
        
    ID <- row.names(filtered_IJC_condition)
        
    IJC1 <- apply(filtered_IJC_condition,1,paste,collapse=",")
        
    SJC1 <- apply(filtered_SJC_condition,1,paste,collapse=",")
        
    IJC2 <- apply(filtered_IJC_control,1,paste,collapse=",")
        
    SJC2<- apply(filtered_SJC_control,1,paste,collapse=",")
        
    rMATS_STAT <- data.frame(ID, IJC1, SJC1, IJC2, SJC2)
        
    rMATS_STAT <- cbind(rMATS_STAT, inclengths)
        
    file4 <- paste("../rmats_stat/hotspot-alldisease_vs_healthybm/",comparison,".txt",sep="")
    print(head(file4))
    write.table(rMATS_STAT, file4, quote=FALSE, row.names=FALSE, col.names=TRUE, sep = "\t")
        
    file5 <- paste("../rmats_stat/hotspot-alldisease_vs_healthybm/",test,".txt",sep="")
    write(condition, file5)
        
    file6 <- paste("../rmats_stat/hotspot-alldisease_vs_healthybm/","_healthy_bm",".txt",sep="")
    write(control, file6)
    } 
}

sapply(colnames(binary_subset), make_rmats_stat_files)

```

##Generating BASH files to submit on HPC

##Generating BATCH script for HPC
\\#!/bin/sh

\\#SBATCH --mail-type=ALL
\\#SBATCH --mail-user=hershbc@ccf.org
\\#SBATCH --job-name=bash5
\\#SBATCH -N1
\\#SBATCH --exclusive

srun python2.7 rMATS_unpaired.py A3SS_MDS_ZRSR2_Other_versus_MDS_SF_WT.txt output/2019_01_28_A3SS_MDS_ZRSR2_Other_versus_MDS_SF_WT_pvalue/ 20 0.1 > output/2019_01_28_A3SS_MDS_ZRSR2_Other_versus_MDS_SF_WT_pvalue/log.txt
```{r}
inputfiles <- list.files("../rmats_stat/hotspot-alldisease_vs_healthybm/", pattern = "versus")
outputfiles <- lapply(inputfiles, function(x) sub(".txt","",x))

makebashscript <- function(inputfile, outputfile){
  command <- paste("python2.7 rMATS_unpaired.py"," ",inputfile," ","output/",outputfile,"_pvalue/"," 20"," 0.05 > ","output/",outputfile,"_pvalue/log.txt", sep="")
  outputfile <- paste("mkdir output/",outputfile,"_pvalue",sep="")
  x <- c(outputfile, command)
  
}


x <- mapply(makebashscript, inputfiles, outputfiles)

#Make several bash files to split up the job on cluster
makebashfiles <- function(gene){

  y <- x[grep(gene,x)]
  job <- paste("#SBATCH --job-name=rmats_stat_",gene,sep="")
  y <- paste("srun ",y)
  #add shebang line
  y <- c("#!/bin/bash","#SBATCH --mail-type=ALL","#SBATCH --mail-user=hershbc@ccf.org",job,"#SBATCH -N1","#SBATCH --exclusive","module load python/2.7.13",y)
  #write out to file
  filename <- paste("../rmats_stat/hotspot-alldisease_vs_healthybm/batchscript_",gene,".sh",sep = "")
  fileConn<-file(filename)
  writeLines(y, fileConn)
  close(fileConn)
}

lapply(list("DDX41","PRPF8"),makebashfiles)
```









