---
title: "rMATS Processing: Subsets rMATS output for rMATS-STAT"
output: html_document
date: "3/13/2019"
output:
  html_document:
    toc: true
    toc_depth: 3
---

This document creates in the input files for rMATS_STAT, the bashscripts joining multiple jobs for the HPC and the batchscripts needed to submit them to the HPC

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(data.table)
```


####Read in final mutations file
```{r}
#somatic mutations
mutations <- read.csv("../inputfiles/filtered_tidied/2019_03_01_combined_mutations_deletions_controls.txt", header = TRUE, sep="\t", stringsAsFactors = FALSE)
binary <- read.csv("../inputfiles/filtered_tidied/2019_03_01_combined_mutations_deletions_controls_binary.txt", header=TRUE, sep="\t", stringsAsFactors = FALSE)
ids <- read.csv("../inputfiles/filtered_tidied/2019_03_01_filtered_inclengths.txt", header = TRUE, sep="\t", stringsAsFactors = FALSE)
length(unique(mutations$array_id))
```

```{r}
#List for disease subtype
AML <- mutations[which(mutations$mkh=="AML"),"array_id"]
MDS <- mutations[which(mutations$mkh=="MDS"),"array_id"]
CMML <- mutations[which(mutations$mkh=="CMML"),"array_id"]
MDS_MPN_RS_T <- mutations[which(mutations$mkh=="MDS/MPN-RS-T"),"array_id"]
MDS_MPN_U <- mutations[which(mutations$mkh=="MDS/MPN-U"),"array_id"]


binary_subset <- binary[,grep("SF3B1|SRSF2|U2AF1",colnames(binary))]

#Combine SF3B1 Hotspot and Missense
binary_subset$SF3B1 <- as.numeric(ifelse(binary_subset$SF3B1_Hotspot==1|binary_subset$SF3B1_missense==1,"1","0"))

#Remove comparisons that I'm not currently interested in.
binary_subset$SF3B1_Hotspot <- NULL
binary_subset$SF3B1_missense <- NULL
binary_subset$SRSF2_missense <- NULL
binary_subset$U2AF1_missense <- NULL

#Adding SF_WT column (don't include healthy bm!)
binary_subset$SF_WT <- ifelse(rowSums(binary_subset)==0 & !rownames(binary_subset) %in% mutations[mutations$mkh=="healthy_bm","array_id"],"1","0")

k <- which(binary_subset>=1, arr.ind=TRUE)
binary_subset[k] <- rownames(k)

binary_subset[binary_subset == 0] <- NA
```

###Read in rMATS output
```{r}
filtered_IJC <- fread("../inputfiles/filtered_tidied/2019_03_01_filtered_includedjunctioncounts.txt", sep="\t", stringsAsFactors = FALSE, data.table=FALSE)
rownames(filtered_IJC) <- filtered_IJC$V1
filtered_IJC$V1 <- NULL

filtered_SJC <- fread("../inputfiles/filtered_tidied/2019_03_01_filtered_skippedjunctioncounts.txt",  sep="\t", stringsAsFactors = FALSE, data.table=FALSE)
rownames(filtered_SJC) <- filtered_SJC$V1
filtered_SJC$V1 <- NULL

inclengths <- fread("../inputfiles/filtered_tidied/2019_03_01_filtered_inclengths.txt", sep="\t", stringsAsFactors = FALSE, data.table=FALSE)
rownames(inclengths) <- inclengths$V1
inclengths$V1 <- NULL
```
###Remove MXE
```{r}
filtered_IJC <- filtered_IJC[which(inclengths$type!="MXE"),]
filtered_SJC <- filtered_SJC[which(inclengths$type!="MXE"),]
inclengths <- inclengths[which(inclengths$type!="MXE"),]
```

##Creating RMATS_STAT input divided by disease type and splicing factor mutation
These SF mutations have enough samples to be run for each disease
```{r}
make_rmats_stat_files <- function(column, disease, disease_subset){

  #disease <- "MDS"
  test <- column
  
  condition <- binary_subset[,column]
  condition <- condition[!is.na(condition)]
  print(condition)
  
  control <- binary_subset$SF_WT[!is.na(binary_subset$SF_WT)]
  
  #Subset for disease
  condition <- condition[condition %in% disease_subset]
  control <- control[control %in% disease_subset]
  
  print(column)
  print(length(condition))
  print(length(control))
  
  if(length(condition) >= 2 & length(control>= 2)){
    
    comparison <- paste(disease,"_",test,"_versus_",disease,"_SF_WT", sep="")   
    
    ##Subset the dataframe for samples of interest
        
    filtered_IJC_condition <- filtered_IJC[,which(colnames(filtered_IJC) %in% condition)]
    filtered_SJC_condition <- filtered_SJC[,which(colnames(filtered_SJC) %in% condition)]
        
    filtered_IJC_control <- filtered_IJC[,which(colnames(filtered_IJC) %in% control)]
    filtered_SJC_control <- filtered_SJC[,which(colnames(filtered_SJC) %in% control)]
        
    ###Create final dataframe
        
    ID <- row.names(filtered_IJC_condition)
        
    IJC1 <- apply(filtered_IJC_condition,1,paste,collapse=",")
        
    SJC1 <- apply(filtered_SJC_condition,1,paste,collapse=",")
        
    IJC2 <- apply(filtered_IJC_control,1,paste,collapse=",")
        
    SJC2<- apply(filtered_SJC_control,1,paste,collapse=",")
        
    rMATS_STAT <- data.frame(ID, IJC1, SJC1, IJC2, SJC2)
        
    rMATS_STAT <- cbind(rMATS_STAT, inclengths)
        
    file4 <- paste("../rmats_stat/hotspot_vs_sfWT/",comparison,".txt",sep="")
    print(head(file4))
    write.table(rMATS_STAT, file4, quote=FALSE, row.names=FALSE, col.names=TRUE, sep = "\t")
        
    file5 <- paste("../rmats_stat/hotspot_vs_sfWT/",disease,"_",test,".txt",sep="")
    write(condition, file5)
        
    file6 <- paste("../rmats_stat/hotspot_vs_sfWT/",disease,"_SF_WT",".txt",sep="")
    write(control, file6)
    } 
}

sapply(colnames(binary_subset), disease="MDS", disease_subset=MDS, make_rmats_stat_files)
sapply(colnames(binary_subset), disease="AML", disease_subset=AML, make_rmats_stat_files)
sapply(colnames(binary_subset), disease="CMML", disease_subset=CMML, make_rmats_stat_files)
sapply(colnames(binary_subset), disease="MDS-MPN-U", disease_subset=MDS_MPN_U, make_rmats_stat_files)
```


##Generating BASH files to submit on HPC

##Generating BATCH script for HPC
\\#!/bin/sh

\\#SBATCH --mail-type=ALL
\\#SBATCH --mail-user=hershbc@ccf.org
\\#SBATCH --job-name=bash5
\\#SBATCH -N1
\\#SBATCH --exclusive

srun python2.7 rMATS_unpaired.py A3SS_MDS_ZRSR2_Other_versus_MDS_SF_WT.txt output/2019_01_28_A3SS_MDS_ZRSR2_Other_versus_MDS_SF_WT_pvalue/ 20 0.1 > output/2019_01_28_A3SS_MDS_ZRSR2_Other_versus_MDS_SF_WT_pvalue/log.txt
```{r}
inputfiles <- list.files("../rmats_stat/hotspot_vs_sfWT/", pattern = "versus")
outputfiles <- lapply(inputfiles, function(x) sub(".txt","",x))

makebashscript <- function(inputfile, outputfile){
  command <- paste("python2.7 rMATS_unpaired.py"," ",inputfile," ","output/",outputfile,"_pvalue/"," 20"," 0.05 > ","output/",outputfile,"_pvalue/log.txt", sep="")
  outputfile <- paste("mkdir output/",outputfile,"_pvalue",sep="")
  x <- c(outputfile, command)
  
}


x <- mapply(makebashscript, inputfiles, outputfiles)

#Make several bash files to split up the job on cluster
makebashfiles <- function(gene){

  y <- x[grep(gene,x)]
  job <- paste("#SBATCH --job-name=rmats_stat_",gene,sep="")
  y <- paste("srun ",y)
  #add shebang line
  y <- c("#!/bin/bash","#SBATCH --mail-type=ALL","#SBATCH --mail-user=hershbc@ccf.org",job,"#SBATCH -N1","#SBATCH --exclusive","module load python/2.7.13",y)
  #write out to file
  filename <- paste("../rmats_stat/hotspot_vs_sfWT/batchscript_",gene,".sh",sep = "")
  fileConn<-file(filename)
  writeLines(y, fileConn)
  close(fileConn)
}

lapply(list("SF3B1","SRSF2","U2AF1"),makebashfiles)
```









