---
title: "Filtering for Biological Significance"
author: "Courtney Hershberger"
date: "1/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(matrixStats)
```

Rick had a good idea yesterday. He said that to remove noise from the rMATS dataset, remove any splicing changes that >10% or <90% in a single sample because they will never be biologically interesting. 
```{r cars}
IJC <- read.table("/Volumes/padgetrlab/courtney/MLL/all/combinedrmatstable/MXE_filteredlowcount_IJC.txt", header=T, stringsAsFactors = FALSE, sep=" ")
SJC <- read.table("/Volumes/padgetrlab/courtney/MLL/all/combinedrmatstable/MXE_filteredlowcount_SJC.txt", header=T, stringsAsFactors = FALSE, sep=" ")
inclengths <- read.table("/Volumes/padgetrlab/courtney/MLL/all/combinedrmatstable/MXE_inclengths.txt", header=T, stringsAsFactors = FALSE, sep=" ")
filteredlowcount <- read.table("/Volumes/padgetrlab/courtney/MLL/all/combinedrmatstable/MXE_filteredlowcount.txt", header=T, stringsAsFactors = FALSE, sep=" ")
```

###Calculate Inclusion Levels
Inclusion Level=(I/lI)/(I/lI+S/lS) 
```{r}  
###Normalize to lengths
#Normalize IJC to length

IJC_matrix <- as.matrix(IJC)
class(IJC_matrix) <- "numeric"
IJC_matrix <- t(IJC_matrix)
normalizedIJC1 <- IJC_matrix/inclengths$IncFormLen[col(IJC_matrix)]

#Normalize SJC to length
SJC_matrix <- as.matrix(SJC)
class(SJC_matrix) <- "numeric"
SJC_matrix <- t(SJC_matrix)
normalizedSJC1 <- SJC_matrix/inclengths$SkipFormLen[col(SJC_matrix)]

#Calculate inclevel for each 
inclevel <- normalizedIJC1/(normalizedIJC1+normalizedSJC1)
```

###Inc Level Spot Check
I ran the above code, subsetting for the first 10 observations and variables then I spot checked the calculation of the inclusion levels by hand. 
```{r comment=''}
cat(readLines('/Volumes/padgetrlab/courtney/MLL/Scripts/rmd_images/2019_01_09_inclevelspotcheck.txt'), sep = '\n')
```


```{r}
inclevel <- t(inclevel)

hist(rowMeans(inclevel))

```
```{r}
inclevel_df <- data.frame(inclevel)
inclevel_df$relevant <- "discard"

inclevel_df[rowMins(inclevel,na.rm = TRUE)>=0.1 | rowMaxs(inclevel,na.rm = TRUE)<=0.9 ,"relevant"] <- "keep"

hist(rowMins(inclevel,na.rm = TRUE))
hist(rowMaxs(inclevel,na.rm = TRUE))

table(inclevel_df$relevant)

ID <- rownames(inclevel_df[inclevel_df$relevant=="keep",])

```

```{r}
IJC_sorted <- IJC[rownames(IJC) %in% ID,]
write.table(IJC_sorted, "/Volumes/padgetrlab/courtney/MLL/all/combinedrmatstable/biologicalrelevance/MXE_filteredlowcount_IJC.txt", col.names=T, quote = FALSE, sep=" ")
SJC_sorted <- SJC[rownames(SJC) %in% ID,]
write.table(SJC_sorted, "/Volumes/padgetrlab/courtney/MLL/all/combinedrmatstable/biologicalrelevance/MXE_filteredlowcount_SJC.txt", col.names=T, quote = FALSE, sep=" ")
inclengths <- inclengths[rownames(inclengths) %in% ID,]
write.table(inclengths, "/Volumes/padgetrlab/courtney/MLL/all/combinedrmatstable/biologicalrelevance/MXE_inclengths.txt", col.names=T, quote = FALSE, sep=" ")
filteredlowcount <- filteredlowcount[rownames(filteredlowcount) %in% ID,]
write.table(filteredlowcount, "/Volumes/padgetrlab/courtney/MLL/all/combinedrmatstable/biologicalrelevance/MXE_filteredlowcount.txt", col.names=T, quote = FALSE, sep=" ")

```