---
title: "Matching Files"
author: "Courtney Hershberger"
date: "3/01/2019"
output:
  html_document:
    toc: true
    toc_depth: 3
---

I have several pieces of data from MLL that have been individually filtered. Today I will also tidy them. 
1) Somatic Mutations information (filtered 2/20/19)
2) Datastatus (age, sex) (no need to filter)
3) CNV information (filtered Nov 8)
3a) Karyotype information (recieved March 7)
4) Expression (unfiltered)
5) Splicing counts(Filtered for biological relevance 1/10/19 then filtered for mis-labeling 2/24/19)

Purpose 1:
For various reasons, these files contain different sets of patient samples. I know that 1289 patients and 63 controls have information for all files. I need to remove patients that do not fall into that category. I will create a single folder ininputfiles/filtered_tidied/ that contains all of these files. At this step samples with CNV data that does not match karyotype data will be removed.

Purpose 2:
The CNV data can be combined with the somatic mutations data. The Datastatus (age, sex) information will also be combined with somatic mutations


Final Files:
1) Somatic/deletion/sex/age
2) Expression
3) rMATS tables(SJC, IJC, inclevels)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(edgeR)
library(data.table)
library(stringr)
library(stats)
library(rstudioapi)
library(ggplot2)
```

```{r}
getwd()
```



##Read in Files
###Somatic Mutations
```{r}
mutations <- read.table("../inputfiles/filtered/2019_02_22_mutations_all_filtered.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)

mutations[which(mutations$mkh=="s-AML"),"mkh"] <- "AML"

#Add RNA VAF
mutations$rna_AD_alt <- as.numeric(mutations$rna_AD_alt)
mutations$rna_AD_ref <- as.numeric(mutations$rna_AD_ref)
mutations$rna_vaf <- mutations$rna_AD_alt/(mutations$rna_AD_ref + mutations$rna_AD_alt)

mutations_orig <- mutations
```

###Datastatus
```{r}
datastatus <- read.table("../inputfiles/filtered/2018_11_01_finalsamplelist.txt", sep = "\t", header = FALSE, stringsAsFactors = FALSE)

datastatus[which(datastatus$V4=="s-AML"),"V4"] <- "AML"
```

###CNV
```{r}
cnv <- read.table("../inputfiles/filtered/deletions_filtered.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)

cnv[which(cnv$Mkh=="s-AML"),"Mkh"] <- "AML"

#bring in karyotype 
karyotype <- read.table("../inputfiles/raw/karyotype.txt", sep = "\t", header=TRUE, na.strings = "NA", stringsAsFactors = FALSE)

cnv <- merge(cnv, karyotype, by="array_id")

#How many X deletions are unconfirmed?
x <- cnv[which(cnv$confirmedx=="unconfirmed" & cnv$chrom=="X"),"array_id"]
#13

#How many 7 deletions are unconfirmed?
seven <- cnv[which(cnv$confirmed7=="unconfirmed" & cnv$chrom=="7"),"array_id"]
#10 

#How many 5 deletions are unconfirmed?
five <- cnv[which(cnv$confirmed5=="unconfirmed" & cnv$chrom=="5"),"array_id"]
#1

#How many 17 deletions are unconfirmed?
seventeen <- cnv[which(cnv$confirmed17=="unconfirmed" & cnv$chrom=="17"),"array_id"]
#4

karytypequestionable <- unique(c(x, seven, five, seventeen))
#26 samples being removed. 

cnv <- cnv[,!colnames(cnv) %in% c("confirmed5","confirmed7","confirmedx","confirmed17")]
deletions<- cnv
```

###Expression
```{r}
mdsexpression <- read.delim("../inputfiles/raw/expression/mds_counts.txt")
healthybmexpression <- read.delim("../inputfiles/raw/expression/Control_cpms.txt")

controls <- colnames(healthybmexpression)

expression <- merge(mdsexpression, healthybmexpression, by=0)
row.names(expression) <- expression[,1]
expression <- expression[,-1]

expressioncol <- colnames(expression)

y <- DGEList(counts=expression)

normalized <- log2(cpm(y))

normalized <- data.frame(normalized)

AML_CMML <- read.delim("../inputfiles/raw/expression/AML_CMML_expr.txt")
MDS_MPNs <- read.delim("../inputfiles/raw/expression/MDS_MPN_expr.txt")

newexpression <- merge(AML_CMML, MDS_MPNs, by=0)
row.names(newexpression) <- newexpression[,1]
newexpression <- newexpression[,-1]

log2cpm <- merge(newexpression, normalized, by=0)
row.names(log2cpm) <- log2cpm[,1]
log2cpm <- log2cpm[,-1]
expression_normalized <- log2cpm
```

###rMATS Inclusion Levels
```{r}
readintables <- function(files){
  filename <- paste("../inputfiles/filtered/rmats/", files, sep="")
  rmats <- fread(filename, stringsAsFactors = FALSE, sep=" ")
  rmats <- as.data.frame(rmats)
  rownames(rmats) <- rmats$V1
  rmats <- rmats[,-1]
  rmats$type <- str_extract(files, "A3SS|A5SS|RI|MXE|SE")
return(rmats)
}
list_rmats <- lapply(list.files("../inputfiles/filtered/rmats/", pattern = "filteredlowcount.txt"), readintables)

RMATS <- do.call(rbind, list_rmats)

rm(list_rmats)

type <- RMATS[,"type", drop=FALSE]
RMATS_Inc <- RMATS[,-length(RMATS)]
```

###rMATS SJC
```{r}
readintables <- function(files){
  filename <- paste("../inputfiles/filtered/rmats/", files, sep="")
  rmats <- fread(filename, stringsAsFactors = FALSE, sep=" ")
  rmats <- as.data.frame(rmats)
  rownames(rmats) <- rmats$V1
  rmats <- rmats[,-1]

return(rmats)
}
list_rmats <- lapply(list.files("../inputfiles/filtered/rmats/", pattern = "SJC"), readintables)

SJC <- do.call(rbind, list_rmats)

```

###rMATS IJC
```{r}
readintables <- function(files){
  filename <- paste("../inputfiles/filtered/rmats/", files, sep="")
  rmats <- fread(filename, stringsAsFactors = FALSE, sep=" ")
  rmats <- as.data.frame(rmats)
  rownames(rmats) <- rmats$V1
  rmats <- rmats[,-1]

return(rmats)
}
list_rmats <- lapply(list.files("../inputfiles/filtered/rmats/", pattern = "IJC"), readintables)

IJC <- do.call(rbind, list_rmats)

rm(list_rmats)
```

###rMATS inclengths
```{r}
readintables <- function(files){
  filename <- paste("../inputfiles/filtered/rmats/", files, sep="")
  rmats <- fread(filename, stringsAsFactors = FALSE, sep=" ")
  rmats <- as.data.frame(rmats)
  rownames(rmats) <- rmats$V1
  rmats$type <- str_extract(files, "A3SS|A5SS|SE|RI|MXE")
  rmats <- rmats[,-1]

return(rmats)
}
list_rmats <- lapply(list.files("../inputfiles/filtered/rmats/", pattern = "inclengths"), readintables)

inclengths <- do.call(rbind, list_rmats)

write.table(inclengths, "../inputfiles/filtered_tidied/2019_03_01_filtered_inclengths.txt", quote = FALSE, sep = "\t")

rm(list_rmats)
```

```{r}
controls_id <- read.table("../inputfiles/filtered/controls_ids.txt", stringsAsFactors = FALSE)
controls_id$V1 <- paste("X",controls_id$V1, sep = "")
```


##Join Mutations/Deletions/Datastatus
###Mutation and Deletion Status of all samples
```{r}
#somatic mutations
mutations <- mutations[,c("array_id","symbol","consequence","mkh","HGVSp","vaf","ref","pos","alt","chrom","rna_vaf")]
mutations$caryotype1 <- NA
mutations$number.cells <- NA
mutations$caryotype2 <- NA
mutations$number.cells.1 <- NA
mutations$caryotype3 <- NA
mutations$number.cells.2 <- NA

#grab everything after the . of mutations$HGVSp 
mutations$protein_change <- gsub("(.+)\\.","",mutations$HGVSp)

###Separate Protein Change that is not NULL
protein_change <- mutations[mutations$protein_change!="NULL",]
protein_changenull <- mutations[mutations$protein_change=="NULL",]


###Convert Amino Acid names from 3 letters to 1
a <- str_match_all(protein_change$protein_change, "[A-Z]{1}[a-z]{2}")

convert <- function(l) {
  map <- c("A", "R", "N", "D", "C", "E", "Q", "G", "H", "I",
           "L", "K", "M", "F", "P", "S", "T", "W", "Y", "V", "*")

  names(map) <- c("Ala", "Arg", "Asn", "Asp", "Cys", "Glu", "Gln",
                  "Gly", "His", "Ile", "Leu", "Lys", "Met", "Phe",
                  "Pro", "Ser", "Thr", "Trp", "Tyr", "Val", "Ter")

  sapply(strsplit(l, "(?<=[A-Z]{3})", perl = TRUE),
         function(x) paste(map[x], collapse = ""))
}

b <- lapply(a, function(x) convert(x))
a <- lapply(a, function(x) as.character(x[,1]))


for(x in 1:length(b)){
names(b[[x]]) <- a[[x]]
}

c <- list()

for(x in 1:length(b)){
c[x] <-  str_replace_all(as.character(protein_change$protein_change[x]), b[[x]])
}

protein_change$protein_change <- c

###Recombine Dataframes
mut_filtered <- rbind(protein_change, protein_changenull)
mut_filtered$protein_change <- as.character(mut_filtered$protein_change)
mut_filtered$HGVSp = NULL

#samples with deletions
deletions <- cnv
deletions$consequence <- "chrom_deletion"
deletions$protein_change <- "chrom_deletion"
names(deletions)[names(deletions) == "Mkh"] <- "mkh"
deletions$vaf <- NA 
deletions$ref <- NA
deletions$pos <- NA
deletions$alt <- NA
deletions$rna_vaf <- NA

deletions <- deletions[,c("array_id","symbol","consequence","mkh","protein_change","vaf","ref","pos","alt","chrom","caryotype1","number.cells","caryotype2","number.cells.1","caryotype3","number.cells.2","rna_vaf")]

#Join deletion data with mutation data
allmutations <- rbind(deletions, mut_filtered)

#Add samples with no mutations but do have disease AND healthy bone marrow controls
healthybm <- data.frame(V1=controls_id$V2, V2=NA, V3=NA, V4="healthy_bm", V5=NA, V6=NA, V7=NA, V8=NA, V9=NA, V10=NA)
datastatus <- rbind(datastatus, healthybm)
dim(datastatus)
names(datastatus)[1] <- "array_id"
names(datastatus)[2] <- "sex"
names(datastatus)[3] <- "age"
names(datastatus)[4] <- "mkh"
dataset_nomut <- unique(datastatus[which(!datastatus$array_id %in% allmutations$array_id),c("array_id","mkh")])
dataset_nomut$symbol <- NA
dataset_nomut$consequence <- NA
dataset_nomut$protein_change <- NA
dataset_nomut$vaf <- NA
dataset_nomut$ref <- NA
dataset_nomut$pos <- NA
dataset_nomut$alt <- NA
dataset_nomut$chrom <- NA
dataset_nomut$caryotype1 <- NA
dataset_nomut$number.cells <- NA
dataset_nomut$caryotype2 <- NA
dataset_nomut$number.cells.1 <- NA
dataset_nomut$caryotype3 <- NA
dataset_nomut$number.cells.2 <- NA
dataset_nomut$rna_vaf <- NA

#Combining to create final dataset
allmutations <- rbind(allmutations, dataset_nomut)

dim(allmutations)
length(unique(allmutations$array_id))

#Adding datastatus (age, sex)
allmutations <- merge(allmutations, datastatus[,c("array_id","sex","age")], by="array_id")
```


##Filtering/Tidying
###Remove samples not present in all files
```{r}
x <- colnames(RMATS_Inc)[colnames(RMATS_Inc) %in% datastatus$array_id]
x <- x[x %in% colnames(expression_normalized)]
y <- controls_id$V2
z <- colnames(RMATS_Inc)[grep("control",colnames(RMATS_Inc))]

rmats_namestokeep <- c(x,z)
length(rmats_namestokeep)
other_namestokeep <- c(x,y)
length(other_namestokeep)
```
**1347 samples remaining**

###Remove samples where cnv doesn't match karyotype
```{r}
rmats_namestokeep <- rmats_namestokeep[!rmats_namestokeep %in% karytypequestionable]
other_namestokeep <- other_namestokeep[!other_namestokeep %in% karytypequestionable]

length(rmats_namestokeep)
length(other_namestokeep)
```

##Write out Filtered/Tidied files
###Somatic/Deletion/Datastatus files
```{r}
allmutations <- allmutations[which(allmutations$array_id %in% other_namestokeep),]

write.table(allmutations, "../inputfiles/filtered_tidied/2019_03_01_combined_mutations_deletions_controls.txt", quote = FALSE, sep = "\t")
```

###RMATS Files
Convert control IDs back to original MLL_IDs, I never should have changed them in the first place. 
```{r}
RMATS_Inc <- RMATS_Inc[,colnames(RMATS_Inc) %in% rmats_namestokeep]
setnames(RMATS_Inc, old=colnames(RMATS_Inc[,controls_id$V1]), new=controls_id$V2)

write.table(RMATS_Inc, "../inputfiles/filtered_tidied/2019_03_01_filtered_inclusionlevels.txt", quote = FALSE, sep = "\t")
rm(RMATS_Inc)

SJC <- SJC[,colnames(SJC) %in% rmats_namestokeep]
setnames(SJC, old=colnames(SJC[,controls_id$V1]), new=controls_id$V2)

write.table(SJC, "../inputfiles/filtered_tidied/2019_03_01_filtered_skippedjunctioncounts.txt", quote = FALSE, sep = "\t")
rm(SJC)

IJC <- IJC[,colnames(IJC) %in% rmats_namestokeep]
setnames(IJC, old=colnames(IJC[,controls_id$V1]), new=controls_id$V2)

write.table(IJC, "../inputfiles/filtered_tidied/2019_03_01_filtered_includedjunctioncounts.txt", quote = FALSE, sep = "\t")
rm(IJC)
```

###Expression
```{r}
expression_normalized <- expression_normalized[,colnames(expression_normalized) %in% other_namestokeep]

dim(expression_normalized)

write.table(expression_normalized, "../inputfiles/filtered_tidied/2019_03_01_expression.txt", quote = FALSE, sep = "\t")
```
Samples missing expression did not contain splicing factor mutations and were removed. 
[1] "MLL_18543" "MLL_11327" "MLL_11335" "MLL_10806" "MLL_10956"


